Optimal (Multiway) Spatial Joins

### 2024-05-27 msj.tex
yufei

    \item {{\bf Type 2:}} either $h$ or $v$ fails to cross at least one rectangle in $\set{r_1, r_2, ..., r_{k-2}}$. Equivalently,
    there exists at least a rectangle $r_i$ (for some $i \in [k-2]$) that covers an endpoint of either $h$ or $v$ or both.
}
Figure~\ref{fig:hv:types} illustrates a result tuple of each type, assuming $k = 4$.

In Section~\ref{sec:hv:type1} (resp., \ref{sec:hv:type2}), we will explain how to produce the result tuples of Type 1 (resp., 2) in the time complexity claimed in Lemma~\ref{lmm:hv}.

\extraspacing {\bf Remark.} In \cite{rdr+11}, Rahul et al.\ studied the problem of storing a set $H$ of horizontal segments and a set $V$ of vertical segments in a data structure such that, given a query rectangle $r$, all the pairs $(h, v) \in H \times V$ satisfying $h \cap v \cap r \ne \emptyset$ can be reported efficiently. They developed a structure of $O(n \log n)$ space that can be built in $O(n \log n)$ time, and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported. Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time. We are unaware of a way to extend their solution to handle H-V $k$-SJ of $k > 3$. Our method for proving Lemma~\ref{lmm:hv} is based on fundamentally different ideas.

\begin{figure}
    \begin{tabular}{cc}
        \includegraphics[height=23mm]{./artwork/type1} &
        \hspace{5mm} \includegraphics[height=27mm]{./artwork/type2} \\
        (a) Type 1 &
        (b) Type 2
    \end{tabular}

@@

As before, let $R_1, ..., R_{k-2}, H$, and $V$ be the input sets of the H-V $k$-SJ problem. Denote by $\J_1$ the set of type-1 result tuples defined in Section~\ref{sec:hv}. In this section, we aim to compute a set $\J^*$ satisfying
\myeqn{
    \J_1
    \subseteq
    \J^*
    \subseteq
    \J(R_1, ..., R_{k-2}, H, V) \label{eqn:hv:type1:J*}
}
where $\J(R_1, ..., R_{k-2}, H, V)$, let us recall, is the join result of the (whole) H-V $k$-SJ. Remember that the output size $\out$ is defined to $|\J(R_1, ..., R_{k-2}, H, V)|$. From $\J^*$, we will report only those $k$-tuples belonging to $\J_1$ and ignore the rest.

@@

Construct
\myeqn{
    R_i' &=& \set{r' \mid r \in R_i \text{ and its trimmed rectangle $r'$ exists}}. \nn
}
Computing the ``segment $h_1$'' for each $r \in R_i$ is an instance of Problem $\mathscr{D}$ (the find-lowest version, with $H$ and $R_i$ as the input). By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r \in R_i$. It thus follows from Section~\ref{sec:bricks} that $R_1', R_2', ..., R_{k-2}'$ can be produced in $O(n \log n)$ total time.

@@

We now solve a $(k-2)$-SJ problem with $R_1', R_2', ..., R_{k-2}'$ as the input using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}). This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R'_1, R'_2, ..., R'_{k-2})$. We prove in Appendix~\ref{app:hv}:

@@

We remind the reader that $\out$ is the output size of the original H-V $k$-SJ problem. As a corollary of Lemma~\ref{lmm:hv:type1:recur-output}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.

@@

We conclude that $\J_1$ --- the set of type-1 result tuples --- can be computed in $F_{k-2}(n, \out) + O(n \log n +  k \cdot \out)$ time.

>>>>

    \item {{\bf Type 2:}} either $h$ or $v$ fails to cross at least one rectangle in $\set{r_1, r_2, ..., r_{k-2}}$. Equivalently,
    there exists at least a rectangle $r_i$ (for some $i \in [k-2]$) that covers an endpoint of either $h$ or $v$ or both.
}
Figure~\ref{fig:hv:types} illustrates a result tuple of each type, assuming $k = 4$. In Section~\ref{sec:hv:type1} (resp., \ref{sec:hv:type2}), we will explain how to produce the result tuples of Type 1 (resp., 2) in the time complexity claimed in Lemma~\ref{lmm:hv}.

\extraspacing {\bf Remark.} In \cite{rdr+11}, Rahul et al.\ studied the problem of storing a set $H$ of horizontal segments and a set $V$ of vertical segments in a data structure such that, given a query rectangle $r$, all the pairs $(h, v) \in H \times V$ satisfying $h \cap v \cap r \ne \emptyset$ can be reported efficiently. They developed a structure of $O(n \log n)$ space that can be built in $O(n \log n)$ time, and can be used to answer a query in $O(\log n + K)$ time, where $n = |H| + |V|$ and $K$ is the number of pairs reported. Their structure can be utilized to solve H-V 3-SJ in $O(n \log n + \out)$ time. We are unaware of a way to extend their solution to handle H-V $k$-SJ of $k > 3$. Our method for proving Lemma~\ref{lmm:hv} is based on fundamentally different ideas even for $k = 3$.

\begin{figure}
    \begin{tabular}{cc}
        \includegraphics[height=27mm]{./artwork/type1} &
        \hspace{5mm} \includegraphics[height=23mm]{./artwork/type2} \\
        (a) Type 1 &
        (b) Type 2
    \end{tabular}

@@

As before, let $R_1, ..., R_{k-2}, H$, and $V$ be the input sets of the H-V $k$-SJ problem. Denote by $\J_1$ the set of type-1 result tuples defined in Section~\ref{sec:hv}. In this section, we aim to compute a set $\J^*$ satisfying
\myeqn{
    \J_1
    \subseteq
    \J^*
    \subseteq
    \J(R_1, ..., R_{k-2}, H, V) \label{eqn:hv:type1:J*}
}
where $\J(R_1, ..., R_{k-2}, H, V)$, let us recall, is the join result of the (whole) H-V $k$-SJ. Remember that the output size $\out$ is defined as $|\J(R_1, ..., R_{k-2}, H, V)|$. From $\J^*$, we will report only those $k$-tuples belonging to $\J_1$ and ignore the rest.

@@

Construct
\myeqn{
    R_i' &=& \set{r' \mid r \in R_i \text{ and its trimmed rectangle $r'$ exists}}. \nn
}
Computing the ``segment $h_1$'' for each $r \in R_i$ is an instance of Problem $\mathscr{D}$ (the find-lowest version, with $H$ and $R_i$ as the input). By symmetry, so is the computation of $h_2, v_1$, and $v_2$ segments for each $r \in R_i$. It thus follows from Section~\ref{sec:bricks} that $R_1', R_2', ..., R_{k-2}'$ can be produced in $O(kn \log n)$ total time.

@@

We now solve a $(k-2)$-SJ problem on the input $\set{R_1', R_2', ..., R_{k-2}'}$ using the algorithm $\A$ supplied (see Lemma~\ref{lmm:hv}). This $(k-2)$-SJ clearly has an input size at most $n$, and let us represent its result as $\J(R'_1, R'_2, ..., R'_{k-2})$. We prove in Appendix~\ref{app:hv}:

@@

As a corollary of Lemma~\ref{lmm:hv:type1:recur-output}, the $(k-2)$-SJ can be settled in $F_{k-2}(n, \out)$ time.

@@

We conclude that $\J_1$ --- the set of type-1 result tuples --- can be computed in $F_{k-2}(n, \out) + O(kn \log n +  k \cdot \out)$ time.
